A user-extensible and adaptable parser architecture

### 2008-08-11 doc/sgai-2008.tex
Dealt with some XXX items.

@@

\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\graphref}[1]{%
    \refwithlabel{fig.}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Figure}{#1}%
}

@@

The framework provides utility and support functions for the actions and
rules.  It provides shared storage to pass data between actions, most
notably the data structures used to represent each connection and mail.  It
loads and validates rules, controls parsing of log files, invokes actions,
tracks how often each rule matches, to optimise rule ordering (see
\sectionref{Rule ordering}), and stores results in the database.  The
framework is independent of the parser being written: the same
functionality will be required for most or all parsers.

The function which finds the rule matching the log line and invokes the
requested action can be expressed in pseudo-code (with indentation denoting
flow of control):

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the varying log lines, and the actions deal with the intricacies
and complications encountered when parsing.  XXX MAYBE REMOVE FIRST TWO
SENTENCES\@.

@@

XXX CARL WONDERS IF THE NUMBERS HERE ARE TYPICAL\@?  ALSO SAY SOMETHING
LIKE REJECTION AND IGNORE DO NOT IMPLY UNGRAMMATICAL\@.

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects.
The parser warns about every unparsed log line, to alert the user that they
need to alter or extend their ruleset; the Postfix parser successfully
parses every input line in the \numberOFlogFILESall{} log files it is
currently being tested with.  The remaining actions have only one or two
associated rules: in some cases this is because there will only ever be one
type of log line for that action, e.g.\ all log lines showing that a remote
client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

Using the \texttt{CONNECT} action as an example, it creates a new data
structure for each client connecting, and saves the data extracted by the
rule (shown in \sectionref{example rule}).  If the data structure already
exists it is treated as a symptom of a bug, and the action issues a warning
containing the full contents of the data structure, plus the log line which
has just been parsed.

@@

This example rule matches the message logged by Postfix when a remote
client connects to deliver mail; it would match the following sample log
line:

@@

% Do not reformat this!
% XXX CONSIDER USING A TABLE ENVIRONMENT; IT DID NOT LOOK GREAT BEFORE.
\begin{tabular}[]{ll}
    \hline
    \noalign{\smallskip}
    Attribute                 & Value                                            \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
    program                   & \daemon{smtpd}                                   \\
    regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
    result\_cols              &                                                  \\
    connection\_cols~\empty{} & client\_hostname = 1; client\_ip = 2             \\
    action                    & CONNECT                                          \\
    priority                  & 0                                                \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
\end{tabular}

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day.  Median log file size is 50MB, containing 285,000 log lines --- large
scale mail servers would have much larger log files.  When generating the
timing data used in this section, \numberOFlogFILES{} log files (totaling
10.08 GB, \numberOFlogLINEShuman{} log lines) were each parsed 10 times and
the parsing times averaged.  Saving results to the database was disabled
for the test runs, as that dominates the run time of the program, and the
tests are aimed at measuring the speed of the parser rather than the speed
of the database and the disks the database is stored on.  The computer used
for test runs was a Dell Optiplex 745:

\begin{tabular}[]{ll}
    \hline
    \noalign{\smallskip}
    XXX         & XXX NEED LABELS\@.                                \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
    CPU         & One dual core 2.40GHz Intel\textregistered{}
                    Core\texttrademark{}2 CPU,                      \\
                & with 32KB L1 cache and 4MB L2 cache.              \\
    RAM         & 2GB 667 MHz DDR RAM\@.                            \\
    Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
\end{tabular}

@@

\newpage{}

>>>>

\newcommand{\tableline}[0]{%
    \noalign{\smallskip}%
    \hline%
    \noalign{\smallskip}%
}

@@

\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\graphref}[1]{%
    \refwithlabel{fig.}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Figure}{#1}%
}
\newcommand{\tableref}[1]{%
    \refwithlabel{tbl.}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  It provides
shared storage to pass data between actions, loads and validates rules,
controls parsing of log files, invokes actions, tracks how often each rule
matches (to optimise rule ordering \sectionref{Rule ordering}), and
stores results in the database.  The framework is largely independent of
the parser being written: the same functionality will be required for most
or all parsers.  The framework is the core of the architecture and is
deliberately quite simple: the rules deal with the varying log lines, and
the actions deal with the intricacies and complications encountered when
parsing.

The function which finds the rule matching the log line and invokes the
requested action can be expressed in pseudo-code as:

@@

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser author will need to write the actions required from scratch,
unless extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, XXX FINISH
THIS SENTENCE\@.  \showgraph{build/graph-action-distribution}{Distribution
of rules per action}{Distribution of rules per action} Unsurprisingly the
action with the highest number of associated rules is \texttt{REJECTION},
the action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(neither \texttt{REJECTION} nor \texttt{IGNORE} imply that the input is
ungrammatical or unparsed).  The parser warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.

@@

Using the \texttt{CONNECT} action as an example, it creates a new data
structure for each client connecting, and saves the data extracted by the
rule.  If the data structure already exists it is treated as a symptom of a
bug, and the action issues a warning containing the full contents of the
data structure, plus the log line which has just been parsed.

@@

The example rule in \tableref{Example rule} matches the message logged by
Postfix when a remote client connects to deliver mail; it would match the
following sample log line:

@@

% Do not reformat this!
\begin{table}[htbp]
    \caption{Example rule}
    \empty{}\label{Example rule}
    \begin{tabular}[]{ll}
        \tableline{}%
        Attribute                 & Value                                            \\
        \tableline{}%
        program                   & \daemon{smtpd}                                   \\
        regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
        result\_cols              &                                                  \\
        connection\_cols~\empty{} & client\_hostname = 1; client\_ip = 2             \\
        action                    & CONNECT                                          \\
        priority                  & 0                                                \\
        \tableline{}%
    \end{tabular}
\end{table}

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day.  Median log file size is 50MB, containing 285,000 log lines --- large
scale mail servers would have much larger log files.  When generating the
timing data used in this section, \numberOFlogFILES{} log files (totaling
10.08 GB, \numberOFlogLINEShuman{} log lines) were each parsed 10 times and
the parsing times averaged.  Saving results to the database was disabled
for the test runs, as that dominates the run time of the program, and the
tests are aimed at measuring the speed of the parser rather than the speed
of the database and the disks the database is stored on.  The computer used
for test runs was a Dell Optiplex 745 described in \tableref{Computer used
to generate statistics}.

\begin{table}[htbp]
    \caption{Computer used to generate statistics}
    \empty{}\label{Computer used to generate statistics}
    \begin{tabular}[]{ll}
        \tableline{}
        Component type  & Component in use                                  \\
        \tableline{}
        CPU             & One dual core 2.40GHz Intel\textregistered{}
                            Core\texttrademark{}2 CPU,                      \\
                        & with 32KB L1 cache and 4MB L2 cache.              \\
        RAM             & 2GB 667 MHz DDR RAM\@.                            \\
        Hard disk       & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
        \tableline{}
    \end{tabular}
\end{table}

@@
